---
title: "US Craft Beer Analysis"
author: "Muthu Palanisamy, Nathan Wall, Nathan Lara"
date: "February 28, 2018"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analysis of US Craft beers Alcohol by Volume (ABV) & Internation Bitterness Unit (IBU)
We will be analyzing data on 2410 US craft beers from 558 different breweries to identify various characteristics of the US craft beer industry. Spefically we will be focusing on a beers ABV and IBU and there geographic properties as well as any relationships to one another.

### Reading & Manipulating Data from source
The data for this analysis is stored on GitHub in the following repository:
https://github.com/nateewall/MSDS6306_CaseStudy1

We will be using R's RCurl package to read this data directly from the web.

```{r Download Dataset}
#if you do not have the RCurl package then install
#install.packages("RCurl")
#Reference RCurl package to read from web
library(RCurl)

# Read Breweries dataset
breweriesURL <- 'https://raw.githubusercontent.com/nateewall/MSDS6306_CaseStudy1/master/Breweries.csv'
Breweries <-read.csv(text=getURL(breweriesURL), header=T)

# Read the Beers Dataset
beersURL <- 'https://raw.githubusercontent.com/nateewall/MSDS6306_CaseStudy1/master/Beers.csv'
Beers <-read.csv(text=getURL(beersURL), header=T)
```


### Exploring the data
Once the data is read into R data.frames from the web we will quickly review the format of the data to make sure all feilds were captured appropriately.

Breweries:

```{r Review Data}
# Review the data
str(Breweries)
str(Beers)
```

Our Breweries data.frame has 558 observations with 551 distinct breweries from 51 states and 384 different cities.

Below we will see how many breweries are located in each of the 51 states in our data:

```{r Statewise Breweries}
require(RColorBrewer)
# How many breweries are present in each state?
BreweryStates <- summary(Breweries$State)
#Print in sorted order
sort(BreweryStates, decreasing = T)
barplot(BreweryStates, 
        xlab="States",
        ylab="Count",
        main = 'Breweries by State',
        col=brewer.pal(12,"Set3"))
```

Colorado has the most breweries in our data with 47 breweries, followed closely by California's 39. Arkansas has the fewest with 2..

### Merge the data sets together to have the geography for each of the beer.

Below is the code used to merge the two data sets by Brewery ID and a quick review of top and the bottom of our new data set. We also join this data to get the full State name for some geographic 

```{r Merge Dataset}
mergedBeerBreweries<-merge(
  Breweries, Beers, by.x = "Brew_ID", by.y ="Brewery_id", all= TRUE
)

#lets include some additional state information for later use
mergedBeerBreweries$State<-trimws(mergedBeerBreweries$State)
# pull the state names to join with abreviated state name
states <- data.frame(cbind(state.abb, tolower(state.name)))
names(states) <- c('abrev','statename') #rename the vars
#join them together to have the complete statename
mergedData<-merge(
  mergedBeerBreweries, states, by.x = "State", by.y ="abrev", all= TRUE
)
#fill in missing statename for DC
mergedData$statename[which(mergedData$State=='DC')] <- "district of columbia"

# Assign appropriate name  
names(mergedData)[2]<-"BreweryName"
names(mergedData)[5]<-"BeerName"

#print the top 6 records
head(mergedData, 6)
#print the bottom 6 records
tail(mergedData, 6)
```

Based on the output above we can see that the data appears to have successfully been merged. Although there are lots of NA's in the for the IBU column so we will perform a count of all NA's in our data.

```{r Find NAs}
# There are multiple ways to get the NA's
# Can go with sapply..
sapply(mergedData, function(x) sum(is.na(x)))
```

It appears the most of the columns have 0 NA's with the exception of 62 NA's for ABV & 1005 NA's for IBU. We are assuming this data was not available for the beers from this group.

### Analyzing the Beers ABV & IBU

First, we will begin by taking a look at the median alcohol content for each of the states to determine if there are any major differences.

```{r ABV & IBU By State}
library(dplyr)
# summarize ABV & IBU by state
plotData <- data.frame(
  mergedData[,c("State","statename","IBU","ABV")] %>%
    group_by(State,statename) %>% 
    summarise_all(funs(median),na.rm = T)
)

plotData <- plotData[order(-plotData$IBU),]
# Set margins 
par(mar=c(5, 6, 4 ,2))

barplot(plotData$IBU, names.arg = plotData$State, 
        xlab="States",
        ylab="Median",
        main = 'IBU by State',
        col=brewer.pal(12,"Set3"))
```

Breweies from Maine (ME) has median IBU of 61ppm isohumulone, 
IBU is a scale that goes from 1-100, standard Budweiser has 7.

```{r Plot ABV}
# sort the data by ABV as this is all about comparing
plotData <- plotData[order(-plotData$ABV),]

# Now plot the data 
barplot(plotData$ABV, names.arg = plotData$State, 
        xlab="States",
        ylab="Median",
        main = 'ABV by State',
        col = brewer.pal(12,"Set2"))
```

We find that our nation's capital holds the tie for the highest Median ABV with a state more commonly associated with whiskey, Kentucky. Both boasting a median ABV of 6.25%. For reference Budweiser comes in aroun 5% ABV.


However, the barplots are little difficult to read so we opted to plot these median values on maps in order to better show the relationship between ABV & IBU we explored these two values geographically.

```{r Geographical Comparision IBU}
library(fiftystater)
library(ggplot2)

plotData$id <- as.character(plotData$statename)
data("fifty_states")
Total <- merge(fifty_states, plotData, by="id")

p <- ggplot(Total, aes(map_id = id)) + 
  # map points to the fifty_states shape data
  geom_map(aes(fill = IBU), map = fifty_states) + 
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())
p1 <- p + scale_fill_continuous(low = "palegreen", high = "darkgreen", guide="colorbar")
# add border boxes to AK/HI
p2 <- p1 + fifty_states_inset_boxes() 
# create title & legend label
p2 + labs(fill = "Median IBU" 
          ,title = "Median IBU for Each State", x="", y="")
```

This shows some of the states with more bitter beer like Maine, Florida, West Virginia, and New Mexico. But also shows the states with less bitter beer like Wisconsin. Now lets see which states move up the ABV Scale.

```{r Geographical Comparision ABV}
q <- ggplot(Total, aes(map_id = id)) + 
  # map points to the fifty_states shape data
  geom_map(aes(fill = ABV), map = fifty_states) + 
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())
q1 <- q + scale_fill_continuous(low = "thistle1", high = "darkred", guide="colorbar")
# add border boxes to AK/HI
q2 <- q1 + fifty_states_inset_boxes() 
q2 + labs(fill = "Median ABV" 
          ,title = "Median ABV for Each State", x="", y="")

```

It is interesting to see the states that seem to favor the High ABV & High IBU like West Virginia & New Mexico, while the bitter Maine falls more towards the middle of ranks in terms of ABV. This may be driven by the state controlled ABV limits.

But, which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?

```{r MaxABV}
#get the max ABV
maxABV <- mergedData[which(mergedData$ABV == max(mergedData$ABV,na.rm = T)), ]
paste0(maxABV$BeerName, " from ", maxABV$City, ", ", maxABV$State, " has the maximum ABV in our data with ", maxABV$ABV, sep=" ")
```

The Lee Hill Series Vol. 5 - Belgian Style Quadrupel Ale from Colorado has the max alcohol by volumne (ABV) beer. However, we have several others that deserve honorable mention.

```{r Max ABV}
top10ABV <- mergedData[order(-mergedData$ABV),][1:10,]
top10ABV[,c("BeerName","City","State","ABV")]
```

Both Colorado & Michigan are representated well in the top 10 with 5 of the top 10 by ABV. 

```{r Max IBU}
#get the max IBU
maxIBU <- mergedData[which(mergedData$IBU == max(mergedData$IBU,na.rm = T)), ]
paste0(maxIBU$BeerName, " from ", maxIBU$City, ", ", maxIBU$State, " has the maximum ABV in our data with ", maxIBU$IBU, sep=" ")
```

Astoria Bitter Bitch Imperial IPA has the max IBU of 138 from Portland, OR. Additionally will we look at the top 10 beers by IBU.

```{r Top10 IBU}
top10IBU <- mergedData[order(-mergedData$IBU),][1:10,]
top10IBU[,c("BeerName","City","State","IBU")]
```

The duplicate Heady Topper's may just an error. Or it may just be so good we felt it needed to be in the data twice. Ranked 4.72/5 on Beer Advocate and #2 overall!


Now before we look directly at the relationship lets get an idea of how IBU & ABV are distributed in our data.

```{r Shiny Histogram ABV & IBU}
# Summary statistics for the ABV variable.
# summary(mergedBeerBreweries$ABV)
library(shiny)
server <- function(input, output) {
  # Define histogram output
  output$beerHist <- renderPlot({  # "beerHist" in UI also
    bc   <- as.numeric(input$var)  # bc = "beer column"
    x    <- mergedData[, bc]
    hist(x, 
         col    = '#8DC13D',  # Green 
         border = 'white',
         main   = paste("Histogram of", names(mergedBeerBreweries[bc]))
    )
  })
}

ui <- fluidPage(
  titlePanel("Charting Beer Data"),
  sidebarLayout(
    sidebarPanel(
      # Dropdown menu for selecting variable from beer data.
      selectInput("var",
                  label = "Select variable",
                  choices = c(
                              "ABV" = 7,
                              "IBU" = 8,
                              "Ounce"=10
                              ),
                  selected = 7)      # Default selection
    ),
    mainPanel(
      plotOutput("beerHist")  # "beerHist" in server also
    )
  )
)

shinyApp(ui = ui, server = server)
#hist(mergedBeerBreweries$ABV)
```

Based on this ABV seems centered around 5.6% with a slight right skew. While IBU's in this dataset appear even more right skewed with a median IBU of 35.


However, lets explore how these two variables relate.
```{r Regression Analysis}
# Is there an apparent relationship between the bitterness 
# of the beer and its alcoholic content? Draw a scatter plot.

plot(mergedData$IBU, mergedData$ABV, 
     xlab="International Bitterness Units",
     ylab="Alcohol by volume",
     main = 'Alcohol Volume Vs. Bitterness correlation',
     col=brewer.pal(2, "Set1"), pch=19)
abline(lm(mergedData$ABV~mergedData$IBU), col="red")
```

There is a postive trend with the linear regression plot suggests there is positive correlation between Alcohol volume and bitterness. However, IBU only explains ~45% of the variability in ABV alone. Perhaps accounting for geography or beer styles would help give us more of an understanding of the bitterness in beer. However, w/ over 100 styles in this data that falls outside the scope of this analysis.
